---
title: "Welcome to my Website!"
author: "**Stringham**"
date: "Utah Valley University"
output: 
    prettydoc::html_pretty:
    theme: cayman
---


```{r setup, include=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(GGally)
library(prettydoc)
library(tidyverse)
library(blogdown)
library(rmarkdown)
library(readr)
library(knitr)
library(DT)
library(modelr)
library(easystats)
library(aRtsy)
library(gifski)
```



# Welcome | [Home](ContactInfo.html)| [About Me](AboutMe.html)| [Resume](Resume.html)| [Project](Project.html)|[New skill](New.html)

# Tying together Sleep and Adiposity, can Sleep Contribute to Adiposity?

 In this research project, I aim to investigate the potential correlation between having an inadequate amount of sleep and adiposity. To achieve this, I have gotten two data sets sourced from the same researcher, who initially sought to understand the sleep patterns and lifestyles prevalent in their area. America is in the top 10 for [obesity](https://data.worldobesity.org/rankings/). With more than 1/3 of Americans getting poor [sleep.](https://www.ncoa.org/adviser/sleep/sleep-statistics/#:~:text=More%20than%20a%20third%20of,38%25%20of%20the%20general%20population.) Along with poor eating habits and lifestyles, the simplest factor of sleep could contribute to weight gain. 
 

*Something to consider before diving in is these are not all the factors that are linked to obesity. 

Allow me to provide an overview of the available data. 

```{r,warnings=FALSE,echo=FALSE,include=FALSE}
dat3 <- read_csv("new_dat3.csv")
```

```{r,echo=FALSE,warning=FALSE}
datatable(dat3, options = list(scrollX = TRUE, scrollY = "300px"))
```

Here is a look at some graphs that show some comparisons.

```{r,include=FALSE}
plot1 <- ggplot(dat3, aes(x = Age, y = WeightKg, color=Trnprt,shape=Gender)) +
  geom_point(alpha = 0.6,size=3) +
  facet_wrap(~Trnprt)+
  labs(title = "Whos doing What?",y="Weight in Kg",color= "Mode of Transportation") +
  theme_minimal()
```

In this plot I wanted to compare what mode of transportation was being used to get an idea of how people were getting around. We can see that most people prefer using an automobile over anything. With a huge proportion of that section being above 100Kgs. 
```{r,echo=FALSE}
plot(plot1)
```

```{r,include=FALSE}
plot3 <- ggplot(dat3, aes(x = WeightKg, y = Age, color = WeightTyp)) +
  geom_point(alpha = 0.7) +
  labs(title = "Weight vs Age", x = "Weight (kg)", y = "Age") +
  theme_minimal()
```

This plot gives a visualization of what members were overweight based on their age, with most overweight people in this study sitting around 20-40 years. 
```{r,echo=FALSE}
plot(plot3)
```

 
This plot shows how many hours of sleep people were getting and how it compared to what kind of weight type they were, overweight being about 100 Kgs. You can see that most of the overweight subjects are staying in the [recommended](https://www.cdc.gov/sleep/features/getting-enough-sleep.html) amount of sleep. This can question if sleep even contributes to being overweight. 
```{r,include=FALSE}
plot2 <- ggplot(dat3,aes(x=Age,y=Sleep_duration,color=WeightTyp))+
  geom_point()+
  facet_wrap(~Gender)+
  labs(title="Hours Slept vs Weight type",y="Hours Slept",color= "Weight Type")+
  theme_minimal()

```

```{r,echo=FALSE}
plot(plot2)
```


This model illustrates the estimated impact on weight for each incremental unit increase in sleep duration and age while keeping other variables constant. The results indicate that for every additional hour of sleep, there is an anticipated decrease in weight of approximately 3.044 Kg. Conversely, age shows a minimal positive effect on weight, suggesting that as individuals age, there may be a slight tendency for weight to increase, albeit to a lesser extent compared to the effect of sleep duration. 
```{r,echo=FALSE}
linear_model <- lm(WeightKg ~ Sleep_duration + Age, data = dat3)
```

```{r,echo=FALSE}
coefficients <- coef(linear_model)
standard_errors <- summary(linear_model)$coefficients[, 2]
p_values <- summary(linear_model)$coefficients[, 4]


summary_df <- data.frame(
  Coefficient = names(coefficients),
  Estimate = coefficients,
  `Std. Error` = standard_errors,
  `Pr(>|t|)` = p_values)

```

```{r,echo=FALSE}
kable(summary_df, align = "c")
```


In our study, sleep duration was how many hours someone slept. But we all know that sleep has different [stages](https://www.sleepfoundation.org/stages-of-sleep). Data was also gathered on what stage of sleep each person had and for how long. With NON-Rem sleep being the best (anything outside the REM stage). 


REM sleep is when dreams are happening, and brain activity looks remarkably like brain activity while you are awake. 


When considering sleep patterns, we can examine the correlation between being overweight and the REM sleep stage. Upon analyzing the graph, it becomes evident that individuals who are overweight exhibit the highest average sleep percentage, indicating that they spend a sizable portion of their sleep in this stage. 
```{r,echo=FALSE}
plot3 <- ggplot(dat3, aes(x = WeightTyp, y = REM_sleep_percentage)) +
  geom_boxplot(fill = "skyblue", color = "blue") +
  labs(title = "Comparison of Weight Category with REM Sleep Percentage",
       x = "Weight Category",
       y = "REM Sleep Percentage") +
  theme_minimal()
```

```{r,echo=FALSE}
plot(plot3)
```

We also possess data regarding alcohol intake before bed, which is known to potentially affect sleep quality for some individuals. With alcohol known to decrease REM sleep early in the night, but as the alcohol wears off, there is a REM sleep rebound, with prolonged REM stages. 
```{r,echo=FALSE,warning=FALSE}
plot4 <- ggplot(dat3, aes(x = Sleep_duration, y = CALC)) +
  geom_boxplot() +
  labs(title = "Relationship Between Sleep Duration and Alcohol consumption before bed",
       x = "Sleep Duration (hours)",
       y = "Alcohol Consumption") +
  theme_minimal()

```

```{r,echo=FALSE}
plot(plot4)
```


Additionally, participants were queried about their engagement in physical activity several times a week. Participants were simply asked to respond with a yes or no. This on paper shows that no physical activity in those tested showed overweight. 
```{r,echo=FALSE}
plot5 <- ggplot(dat3, aes(x = WeightTyp, fill = factor(FAF))) +
  geom_bar(position = "dodge", color = "black") +
  labs(title = "Comparison of Physical Activity with Weight Category",
       x = "Weight Category",
       y = "Count",
       fill = "Physical Activity") +
  theme_minimal()
```

```{r,echo=FALSE}
plot(plot5)
```

As we conclude, it is important to note the significant impact of alcohol on sleep patterns. Disturbances in sleep duration and quality can, in turn, influence overall health, potentially contributing to weight gain in certain instances. However, it is essential to recognize that these are just a few of the numerous factors that may influence obesity, underscoring the need for comprehensive testing and analysis. 





